**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-01-28 Schloss_PacBio16S_PeerJ_2016.Rmd
Trying to get write.paper to run

***Identifying correlates of increased sequencing error.***
In contrast to the 454 and Illumina-based platforms where the sequencing
quality decays with length, the consensus sequencing approach employed by the
PacBio sequencer is thought to generate a uniform distribution of errors. This
makes it impossible to simply trim sequences to high quality regions. Therefore,
we sought to identify characteristics within sequences that would allow us to
identify and remove those sequences with errors using three different
approaches. First, we hypothesized that errors in the barcode and primer would
be correlated with the error rate for the entire sequence. We observed a strong
relationship between the number of mismatches to the barcodes and primers and the
error rate of the rest of the sequence fragment (Figure 1B). Although
allowing no mismatches to the barcodes and primers yielded the lowest error rate,
that stringent criterion removed a large fraction of the reads from the dataset.
Allowing at most one mismatch only marginally increased the error rate while
retaining more sequences in the dataset (Figure 2). Second, we hypothesized that
increased sequencing coverage should yield lower error rates. We found that once
we had obtained 10-fold coverage of the fragments, the error rate did not change
appreciably (Figure 1C). When we compared the error rates of reads with at least
10-fold coverage to those with less coverage, we reduced the error rate by
`r paste(format(range(coverage_reduction), digits=3, nsmall=1), collapse=" to ")`%
(Figure 2). Third, based on the observed correlation between the
predicted and observed error rates, we sought to identify a minimum predicted
error rate that would allow us to reduce the observed error rate. The average
observed error rate for sequences with predicted error rates between 0.01 and
0.10% was linear (Figure 1D). We decided to use a threshold of 0.01% because a
large number of sequence reads were lost when we used a smaller threshold. When
we used this threshold, we were able to reduce the error rate by
`r paste(format(range(pred_error_reduction), digits=3, nsmall=1), collapse=" to ")`%
(Figure 2A). Finally, we quantified the effect of combining filters. We found that
any combination of filters that included the predicted error rate threshold had
the most significant impact on reducing the observed error rate. Furthermore,
the inclusion of the mismatch and coverage filters had a negligible impact on
error rates, but had a significant impact on the number of sequences included in
the analysis. For instance, among the V1-V9 data, requiring sequences to have a
predicted error rate less than 0.01% resulted in a `r format(base_error_reduction["pred_error_reduction","V1V9"], digits=3, nsmall=1)`% and resulted in the removal of `r format(base_nseqs_reduction["pred_error_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences. Adding the mismatch or coverage filter
had no effect on the reduction of error, but resulted in the removal of
`r format(base_nseqs_reduction["oligos_pred_nseqs_reduction","V1V9"], digits=3, nsmall=1)` and
`r format(base_nseqs_reduction["coverage_pred_nseqs_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences, respectively. Use of all filters had no impact on the reduction in the observed
error rate, but resulted in the removal of `r format(base_nseqs_reduction["all_filters_nseqs_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences. The remainder of this paper only uses sequences with a predicted error rate less than 0.01%.

>>>>

***Identifying correlates of increased sequencing error.***
In contrast to the 454 and Illumina-based platforms where the sequencing
quality decays with length, the consensus sequencing approach employed by the
PacBio sequencer is thought to generate a uniform distribution of errors. This
makes it impossible to simply trim sequences to high quality regions. Therefore,
we sought to identify characteristics within sequences that would allow us to
identify and remove those sequences with errors using three different
approaches. First, we hypothesized that errors in the barcode and primer would
be correlated with the error rate for the entire sequence. We observed a strong
relationship between the number of mismatches to the barcodes and primers and the
error rate of the rest of the sequence fragment (Figure 1B). Although
allowing no mismatches to the barcodes and primers yielded the lowest error rate,
that stringent criterion removed a large fraction of the reads from the dataset.
Allowing at most one mismatch only marginally increased the error rate while
retaining more sequences in the dataset (Figure 2). Second, we hypothesized that
increased sequencing coverage should yield lower error rates. We found that once
we had obtained 10-fold coverage of the fragments, the error rate did not change
appreciably (Figure 1C). When we compared the error rates of reads with at least
10-fold coverage to those with less coverage, we reduced the error rate by
`r paste(format(range(base_error_reduction), digits=3, nsmall=1), collapse=" to ")`%
(Figure 2). Third, based on the observed correlation between the
predicted and observed error rates, we sought to identify a minimum predicted
error rate that would allow us to reduce the observed error rate. The average
observed error rate for sequences with predicted error rates between 0.01 and
0.10% was linear (Figure 1D). We decided to use a threshold of 0.01% because a
large number of sequence reads were lost when we used a smaller threshold. When
we used this threshold, we were able to reduce the error rate by
`r paste(format(range(pred_error_reduction), digits=3, nsmall=1), collapse=" to ")`%
(Figure 2A). Finally, we quantified the effect of combining filters. We found that
any combination of filters that included the predicted error rate threshold had
the most significant impact on reducing the observed error rate. Furthermore,
the inclusion of the mismatch and coverage filters had a negligible impact on
error rates, but had a significant impact on the number of sequences included in
the analysis. For instance, among the V1-V9 data, requiring sequences to have a
predicted error rate less than 0.01% resulted in a `r format(base_error_reduction["pred_error_reduction","V1V9"], digits=3, nsmall=1)`% and resulted in the removal of `r format(base_nseqs_reduction["pred_error_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences. Adding the mismatch or coverage filter
had no effect on the reduction of error, but resulted in the removal of
`r format(base_nseqs_reduction["oligos_pred_nseqs_reduction","V1V9"], digits=3, nsmall=1)` and
`r format(base_nseqs_reduction["coverage_pred_nseqs_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences, respectively. Use of all filters had no impact on the reduction in the observed
error rate, but resulted in the removal of `r format(base_nseqs_reduction["all_filters_nseqs_reduction","V1V9"], digits=3, nsmall=1)`% of the sequences. The remainder of this paper only uses sequences with a predicted error rate less than 0.01%.
